\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{times}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}

\title{When Truth Is Retrieved but Ignored: Benchmarking Evidence-Present Indirect Injection in Multi-Document Retrieval-Augmented Generation}
\author{Anonymous}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Retrieval-Augmented Generation (RAG) grounds answers in retrieved documents; indirect prompt injection can appear in retrieved text. We study \emph{evidence-present} settings where truthful supporting documents are already in the retrieved set, while another passage contains instruction-like or manipulative content. We describe a reproducible benchmark construction pipeline over Natural Questions (via KILT) and HotpotQA, a failure decomposition (retrieval vs.\ exposure vs.\ obedience), and evaluate TRIM (Trust-Role Isolation Middleware) against prompt-only baselines. Tabular results are produced from the pilot run logs (e.g.\ \texttt{logs/pilot\_summary.json}) and exported to LaTeX.
\end{abstract}

\section{Introduction}
Large language models augmented with retrieval can still be manipulated by untrusted passages co-retrieved with legitimate evidence. Indirect prompt injection via untrusted content is now well documented in LLM-integrated applications~\cite{greshake2023indirect}, and end-to-end RAG-style pipelines have been studied under systematic manipulations~\cite{destefano2024ragnroll}. Unlike settings where retrieval misses gold evidence, we isolate cases where gold remains accessible yet models obey conflicting directives.

\section{Threat Model}
We consider a multi-document QA RAG pipeline with a static per-query corpus containing gold passages, distractors, and one injected ``poison'' document designed to resemble benign corpus text.

\section{Method}
\subsection{Benchmark construction}
Controlled pools are built per query with gold and distractor documents; a poison document is synthesized from one of nine templates spanning three attack families (answer hijack, false abstention, instruction override) with three templates per family. To avoid toy-benchmark failure modes, every template is framed as a realistic carrier document---a knowledge-base release note, an editorial corrigendum, a data-steward advisory, an internal source-policy memo, a legal redaction notice, a compliance review hold, a help-center FAQ, an approved assistant runbook, or a vendor integration spec---with the adversarial directive embedded inside the carrier's natural section structure rather than stated as a bare imperative. The template is sampled uniformly per question so that variance across templates can be reported.

\subsection{TRIM}
TRIM wraps retrieved content in explicit evidence delimiters with source metadata and instructs the model to treat retrieved text as untrusted evidence, not as instruction authority. The \texttt{trim\_mask} variant additionally replaces directive-like spans inside poison-role passages with a \texttt{[MASKED]} placeholder before prompt assembly; matches are logged per row so that mask-hit rate can be reported alongside defense effectiveness.

\section{Experiments}

\subsection{Setup}
Unless noted otherwise, we use the defaults in \texttt{configs/pilot.yaml}: Natural Questions via KILT and HotpotQA-style items, BM25 retrieval with $k=5$, six distractor passages per pool, and poison ranks drawn from $\{1,5\}$ for the main pilot (and $\{1,3,5\}$ for the ablation script). Generation uses \texttt{Qwen/Qwen2.5-7B-Instruct} with greedy decoding (\texttt{temperature}$=0$) and up to 256 new tokens, either through an OpenAI-compatible server or local Hugging Face weights (\texttt{--hf-fallback}). Each logged row is one full prompt--answer call for a given query, poison placement, and defense.

\subsection{Retrieval sanity (no LLM)}
Before generation experiments, we ran a clean retrieval sanity check on $n=20$ pilot examples using \texttt{scripts/run\_clean\_sanity.py} with BM25 and dense encoders (\texttt{--dense}; \texttt{BAAI/bge-large-en-v1.5}). On this slice, gold evidence appeared in top-$k$ for every example under both sparse and dense retrieval (\texttt{bm25\_recall\_at\_k}$=1.0$, \texttt{dense\_recall\_at\_k}$=1.0$; see \texttt{logs/clean\_sanity\_summary.json}). This supports treating downstream failures as primarily exposure or obedience effects under evidence-present pools, not systematic retrieval miss on the pilot slice.

\subsection{Metrics}
\textbf{ASR (rules)} is the fraction of model outputs flagged by lightweight, attack-family-specific rules as satisfying the adversarial objective (e.g., emitting a malicious target string). \textbf{EM} is exact match of the model answer against benchmark gold strings after normalization. The pilot runner also records token-overlap F1 against gold answers in the JSON logs; Table~\ref{tab:pilot} focuses on ASR and EM for space. Lower ASR is better; higher EM is better.

\subsection{Main pilot}
We run \texttt{scripts/run\_pilot.py} over 100 KILT NQ and 100 Hotpot-style training items (seed~42), sampling a poison template per question from the full nine-template pool, then enumerating both configured poison ranks and all five defenses (\texttt{none}, \texttt{reminder}, \texttt{boundary}, \texttt{trim}, \texttt{trim\_mask}). That yields 200 queries $\times$ 2 ranks $\times$ 5 defenses $=2000$ generation rows; aggregates appear in \texttt{logs/pilot\_summary.json}. Table~\ref{tab:pilot} reports per-defense means exported via \texttt{python scripts/export\_tables.py}. In this run, \texttt{poison\_in\_topk} holds for all rows by construction, and \texttt{gold\_in\_topk} holds for $90.0\%$ of rows (we report aggregates without conditioning for space; logs retain per-row conditions).

Under attack, the undefended baseline reaches ASR $0.363$, confirming the evidence-present phenomenon: directives embedded in carrier documents change answers even when gold is still in top-$k$. Notably, the soft \texttt{reminder} defense slightly \emph{raises} ASR ($0.408$) over the undefended baseline; we read this as the reminder priming the model to attend more strongly to the directive language it was warned about, an unintended interaction prior workshop-scale studies do not appear to have surfaced. The heavier \texttt{boundary} prompt brings ASR back down to roughly the baseline ($0.365$) but at a clear utility cost (EM drops to $0.233$). Structural isolation (\texttt{trim}) lowers ASR to $0.248$ while preserving EM, and masking directive spans (\texttt{trim\_mask}) drives ASR to $0.025$ with a mean of $8.98$ mask hits per poison passage (logged as \texttt{mask\_hits.mean\_hits\_trim\_mask} in the summary). \texttt{trim\_mask} is the only defense that simultaneously achieves the lowest ASR and the highest EM under attack ($0.333$), empirically separating it from both prompt-only wrappers and from the unmasked structural isolation.

\paragraph{Per-template variance.} Because the template is sampled per question, we can report variance across templates from a single pilot run (Table~\ref{tab:pilot-template}). Across the nine templates, undefended ASR ranges from $0.237$ for the relatively soft \texttt{policy\_memo\_uncertainty} carrier to $0.536$ for the imperative \texttt{data\_steward\_advisory} carrier---a $2.3\times$ spread within a single attack family in some cases. The phenomenon is therefore not a single-template artifact: every template induces a non-trivial undefended ASR, but realism and explicit ``output verbatim'' framing matter materially for attack strength.

\paragraph{Attack family and poison rank.} Per-family and per-rank breakdowns are given in Tables~\ref{tab:pilot-family} and~\ref{tab:pilot-rank}. The \texttt{answer\_hijack} and \texttt{instruction\_override} families both produce undefended ASR around $0.39$, indicating that getting the model to emit an attacker-chosen string is roughly as easy as getting it to abandon a known answer; \texttt{false\_abstention} is harder for the attacker (undefended ASR $0.289$), which we read as the model being mildly biased toward producing some answer when gold is visible in context. Attack success is sharply non-monotone in rank: placing the poison at rank~5 yields ASR $0.395$, more than $2.3\times$ the rank~1 ASR of $0.168$, supporting a recency-of-directive reading where the model weights instruction-like spans encountered later in the retrieved context more heavily.

\paragraph{Matched clean-utility control.} To separate attack-induced utility loss from defense-induced utility loss, we ran \texttt{scripts/run\_pilot.py --clean-control}, which rebuilds each pool without poison and re-runs every defense once per question (see \texttt{logs/pilot\_clean\_summary.json}; Table~\ref{tab:pilot-clean}). Refusal rates remain at or below $1\%$ across all defenses, so none of the methods ``wins'' by over-abstention. On poison-free pools, \texttt{trim} and \texttt{trim\_mask} achieve the highest EM and F1 of any defense; in contrast, \texttt{boundary} reduces EM to $0.195$, confirming that long boundary-policy prompts can hurt grounded answering even in the absence of attack.

\subsection{Defense ablation}
We run \texttt{scripts/run\_ablation.py} on a smaller slice controlled by \texttt{--subset} (half NQ, half Hotpot-style), holding the poison template fixed to the first template family, sweeping defenses \texttt{none}, \texttt{reminder}, \texttt{trim}, and \texttt{trim\_mask} across poison ranks $\{1,3,5\}$. Table~\ref{tab:ablation} shows per-mode ASR and EM for the summary currently checked into \texttt{logs/ablation\_summary.json} and exported with the same script as the pilot table.

\begin{table}[t]
\centering
\caption{Pilot metrics by defense under attack (from \texttt{logs/pilot\_summary.json}; regenerate via \texttt{python scripts/export\_tables.py}). $n{=}400$ rows per defense, aggregating over 200 queries, two poison ranks, and a per-question uniform sample over the nine attack templates.}
\label{tab:pilot}
\input{tables/pilot_metrics.tex}
\end{table}

\begin{table}[t]
\centering
\caption{Pilot attack success and EM by attack family (from \texttt{logs/pilot\_summary.json}).}
\label{tab:pilot-family}
\input{tables/pilot_by_family.tex}
\end{table}

\begin{table}[t]
\centering
\caption{Pilot attack success and EM by poison rank (from \texttt{logs/pilot\_summary.json}).}
\label{tab:pilot-rank}
\input{tables/pilot_by_rank.tex}
\end{table}

\begin{table}[t]
\centering
\caption{Matched clean-utility control: EM, F1, and refusal rate on poison-free pools (from \texttt{logs/pilot\_clean\_summary.json}; $n{=}200$ rows per defense).}
\label{tab:pilot-clean}
\input{tables/pilot_clean_metrics.tex}
\end{table}

\begin{table}[t]
\centering
\caption{Ablation metrics by defense mode (from \texttt{logs/ablation\_summary.json}; regenerate via \texttt{python scripts/export\_tables.py}).}
\label{tab:ablation}
\input{tables/ablation_metrics.tex}
\end{table}

\begin{table*}[t]
\centering
\caption{Per-template breakdown on the main pilot (from \texttt{logs/pilot\_summary.json}). ``ASR (none)'' is the undefended baseline ASR for that template; ``ASR (all)'' averages across all five defenses. Templates are grouped by family and sorted within family by descending undefended ASR.}
\label{tab:pilot-template}
\input{tables/pilot_by_template.tex}
\end{table*}

\section{Limitations}
Attack success is measured with rules that approximate human judgment; they can miss subtle successes or over-flag borderline phrasing, so ASR should be read as a tractable proxy rather than ground truth. Poison passages are synthetic and drawn from nine templates spanning three carrier-genre clusters; this is large enough to surface meaningful per-template variance (Table~\ref{tab:pilot-template}) but does not cover the full diversity of open-web carrier text, so absolute ASR magnitudes should be read as a sample over a reproducible but bounded template space. An earlier version of the corpus builder seeded the poison header with the first available ``distractor'' passage; on KILT items with few native distractors, that slot could be padded from a non-primary gold passage, mildly inflating EM for non-masking defenses under attack. The current pipeline restricts the poison header to true distractor text (or omits it entirely) and only pads the surrounding pool, not the poison passage; numbers reported here are from runs after this fix. Natural Questions rows in the KILT-style path may use short surrogate passages when full Wikipedia text is not wired in, which can shift retrieval behavior relative to full-document corpora. Finally, all numbers here depend on the specific base model, decoding parameters, and retrieval backend; changing any of these can move ASR and EM materially without altering the benchmark code.

\section{Ethics}
Adversarial passages are synthetic and used for controlled evaluation in a research setting.

\bibliographystyle{plain}
\bibliography{references}

\end{document}
